% =============================================================
% §1 Introduction
% =============================================================
\section{Introduction}
\label{sec:intro}

Generative AI workloads, dominated by large language model (LLM)
inference, have become the fastest-growing class of compute in
modern data centers. The energy and carbon cost of training and
serving such models has grown by orders of magnitude over the
past five years~\cite{patterson2021carbon}, and inference now
consumes more aggregate FLOPs than training across most production
deployments~\cite{patel2024splitwise}. As LLM-powered services
move from research demos into latency-sensitive applications---voice
assistants, real-time coding assistants, streaming summarization---the
systems community has begun rethinking how inference workloads should
be \emph{scheduled} across heterogeneous compute pools.

This paper studies \textbf{cloud-edge collaborative LLM inference
scheduling}, a deployment paradigm gaining traction for three
reasons: (i)~\textbf{cost}, by serving common-case queries on
cheaper edge accelerators while reserving cloud GPUs for tail
requests; (ii)~\textbf{privacy}, by keeping prompt data on-premises
whenever possible; and (iii)~\textbf{tail latency}, by amortizing
first-token latency through edge proximity. The scheduling
problem---deciding which server should serve each request---now
becomes a critical performance lever, distinct from intra-server
batching engines such as
vLLM~\cite{kwon2023vllm}, Orca~\cite{yu2022orca}, and
Sarathi-Serve~\cite{agrawal2024sarathi}.

\subsection{What makes LLM inference scheduling different}

Existing cluster schedulers (HEFT, GA, PSO, and learning-based
variants like Decima~\cite{mao2019decima} and
RLScheduler~\cite{zhang2020rlscheduler}) were designed for generic
DAG workloads such as Spark jobs or HPC batch tasks. They do not
model the five physical characteristics that fundamentally
distinguish LLM inference:

\begin{enumerate}
\item \textbf{Model weight residency.} LLM weights (14--70~GB
  for LLaMA-7B/13B/70B at FP16/INT8) live persistently in GPU
  memory. Switching models incurs cold-load cost of 5--60~seconds---orders
  of magnitude larger than any per-request computation.

\item \textbf{KV-cache locality.} Each request's KV cache---built
  during prefill, consumed during decode---is bound to the GPU where
  prefill ran. Migrating KV across servers costs hundreds of MB
  of data movement per request, often dominating decode latency.

\item \textbf{Continuous batching.} vLLM-style iteration-level
  batching serves $N$ same-model requests in approximately
  $(1 + (N-1)\cdot \alpha)\cdot T_{\text{solo}}$ time, providing
  throughput speedups of $5$--$8\times$. Crucially, this benefit
  only materializes when the scheduler co-locates same-model
  requests on the same server.

\item \textbf{Memory-bandwidth-bound decode.} Unlike prefill
  (compute-bound), decode is bound by HBM memory bandwidth.
  Decode latency per token is approximately 20~ms on A100-class
  GPUs regardless of compute capacity, meaning that ``throw more
  compute'' does not accelerate decode beyond this floor.

\item \textbf{Two-phase request lifecycle.} Each inference
  request decomposes into (Prefill $\rightarrow$ Decode) with
  strict dependency and shared KV state. This violates the
  assumption of independent stages baked into most DAG
  schedulers.
\end{enumerate}

The combination of these five physics creates scheduling
trade-offs absent from generic workloads: e.g., consolidating
same-model requests on one server (for batching) versus
spreading to avoid contention; placing decode on a
fast-but-distant server (low compute time) versus its sibling
prefill server (no KV migration). \textbf{A scheduler that
ignores AIGC physics cannot exploit these trade-offs}, leaving
substantial efficiency on the table.

\subsection{Why na\"ive approaches fail}

A natural hypothesis is that an off-the-shelf RL scheduler with
sufficient training would \emph{implicitly} learn AIGC-aware
behavior from base features (compute capacity, queue length,
memory utilization). We initially set out to test exactly this
hypothesis.

Our empirical journey reveals a more subtle picture
(Fig.~\ref{fig:pareto}):

\begin{itemize}
\item Generic load-balancing baselines (LeastLoaded,
  ShortestQueue), optimal-DAG heuristics (HEFT, GA, PSO), and
  even modern DRL baselines (A3C-R2N2~\cite{tuli2022a3cr2n} and
  a Decima-inspired GNN variant we
  implemented~\cite{mao2019decima}) all cluster in the same
  Pareto region: \slo{} attainment around 21--24\% and energy
  efficiency around 2.8--3.0 J/token.

\item Our AIGC-aware PPO scheduler, sharing the actor-critic
  family and equivalent pretraining budget with A3C-R2N2 but
  with explicit AIGC state features and reward shaping, achieves
  a \textbf{distinctly better Pareto point}: \slo{} 30.2\%
  ($+28\%$ relative) and energy 2.61~J/token ($-7\%$ relative),
  with $p<0.05$ on \slo{} against all four strongest baselines
  and on energy per token against three of four (PSO is
  directionally consistent but $p=0.11$).

\item However, our 12-component ablation
  (Table~\ref{tab:ablation}, Fig.~\ref{fig:ablation}) reveals a
  paradox: removing any single AIGC component (warm reward,
  batch reward, affinity reward, AIGC state features, even all
  of them jointly) leaves performance essentially unchanged.
  Only continuous batching simulation and adequate pretraining
  stand out as significant components.
\end{itemize}

\subsection{Reconciliation and contributions}

These observations are reconciled by recognizing that
\textbf{AIGC-aware scheduling is a holistic system contribution,
not a clever reward component}. Once a scheduler has
(i)~continuous batching physics in its environment,
(ii)~adequate pretraining samples, and (iii)~some AIGC
awareness in either state or reward, the PPO backbone extracts
equivalent signal from any of the AIGC priors---they are
mutually substitutable in practice. The Pareto improvement is
real, but it comes from the \emph{combination}, not from
individual reward tricks.

\medskip
\noindent This paper makes four contributions:

\smallskip
\noindent\textbf{(C1) An AIGC inference simulation platform}
(M1--M4 milestones) that models the five physical characteristics
described above: model weight residency with LRU eviction (M1),
two-phase prefill/decode tasks with KV-cache memory accounting
(M2), admission-time continuous batching (M3), and
memory-bandwidth floors calibrated to vLLM measurements (M4).
The platform is open-sourced with reproducibility manifests.

\smallskip
\noindent\textbf{(C2) An AIGC-aware PPO scheduler} (RL) that
exposes AIGC physics to the policy via state features
(loaded-model indicator, batch occupancy, sibling-server
affinity, KV-cache size) and reward shaping (warm bonus, batch
bonus, affinity bonus, cloud-overuse penalty). The scheduler
achieves \textbf{Pareto dominance} over eight strong baselines
in \slo{} attainment and energy efficiency.

\smallskip
\noindent\textbf{(C3) A systematic 12-component ablation study}
revealing which AIGC abstractions actually contribute. The
striking null result---that no single AIGC reward or state
component is individually significant---suggests prior works
claiming specific AIGC reward tricks should ablate aggressively
before attributing gains to specific designs.

\smallskip
\noindent\textbf{(C4) An empirical ``regime'' characterization}
showing that AIGC-aware RL achieves strict Pareto dominance only
in the \textbf{moderate-load operating regime}
($\lambda \in [1, 2]$ req/s with uniform model mix), while
maintaining a robust 5--11\% energy efficiency advantage across
all 30 tested configurations. This nuanced finding contradicts
simpler ``always wins'' claims common in DRL scheduling
literature and provides actionable guidance for AIGC scheduling
deployment.

\medskip
The rest of this paper is organized as follows: \S\ref{sec:background}
reviews LLM inference and cloud-edge scheduling background;
\S\ref{sec:formulation} formalizes the scheduling problem with the
five AIGC physics; \S\ref{sec:system} details our simulator design
(\S\ref{sec:simulator}) and AIGC-aware RL scheduler
(\S\ref{sec:rl}); \S\ref{sec:eval} reports our experimental
evaluation; \S\ref{sec:discuss} discusses limitations and
implications; \S\ref{sec:related} surveys related work;
\S\ref{sec:conclusion} concludes.
