% =============================================================
% §2 Background
% =============================================================
\section{Background}
\label{sec:background}

This section provides the background needed to understand the
AIGC physics introduced in \S\ref{sec:intro} and the system
design in \S\ref{sec:system}. We first sketch LLM inference
(\S\ref{sec:bg-inference}) and the continuous-batching serving
model (\S\ref{sec:bg-batching}), then describe cloud-edge
collaborative deployment (\S\ref{sec:bg-cloudedge}), and finally
summarize the scheduling approaches our work builds on or
contrasts with (\S\ref{sec:bg-sched}).

\subsection{LLM Inference: Prefill, Decode, KV Cache}
\label{sec:bg-inference}

Modern LLMs are decoder-only
Transformers~\cite{vaswani2017attention,brown2020gpt3} producing
one output token at a time given a textual prompt. Inference
proceeds in two distinct phases that have markedly different
hardware characteristics.

\textbf{Prefill} is invoked once per request: it ingests the
entire prompt (of length $L_p$ tokens) through a forward pass
in parallel, producing the first output token and---critically---an
intermediate \emph{key-value cache} (KV cache) for each
Transformer layer. Prefill compute scales as $O(L_p \cdot d)$
multiply-accumulate operations and saturates GPU compute
utilization for prompts longer than a few dozen tokens.

\textbf{Decode} is then invoked iteratively, once per output
token, to generate the remaining $L_o$ tokens. Each decode step
re-uses the stored KV cache plus one new token's worth of input,
producing one new token and extending the KV cache by one entry
per layer. Each decode step is fast ($\sim$10--80~ms on
A100-class GPUs), but the \emph{sequence} of $L_o$ steps
dominates end-to-end request latency for long generations.

The KV cache is large ($\approx$0.5--2.5~MB per token per
Transformer layer for LLaMA-class models) and grows
monotonically with each decode step. For a 1000-token
conversation on LLaMA-13B, the KV cache reaches roughly
800~MB---often larger than the activation memory of a single
forward pass. \textbf{Critically, the KV cache is bound to the
GPU where prefill executed}; serving the corresponding decode
steps on a different GPU requires either copying hundreds of
MB across the network (expensive) or re-running prefill
(wasteful).

Beyond the standard Transformer LLMs that are our focus, AIGC
workloads also include text-to-image diffusion models (Stable
Diffusion XL) whose inference pattern---iterative denoising
steps invoking the same U-Net---is structurally different but
shares the model-residency and batching characteristics that
motivate our work.

\subsection{Continuous Batching and Modern Serving Systems}
\label{sec:bg-batching}

Throughput-oriented LLM serving systems exploit a key
optimization: \textbf{batching same-model requests into a
shared forward pass}. With $B$ requests in a batch, each decode
step takes $T_{\text{solo}} \cdot (1 + (B-1) \cdot \alpha)$
wall-clock time rather than $B \cdot T_{\text{solo}}$, providing
a $5$--$8\times$ throughput speedup at modest per-request
latency cost. The decode-phase batching overhead is typically
$<5\%$ per added request because decode is memory-bandwidth-bound;
the prefill-phase overhead is higher ($\sim$30\%) because
prefill is compute-bound and harder to fuse.

The breakthrough enabling efficient batching at scale is
\textbf{continuous (iteration-level) batching}, introduced by
Orca~\cite{yu2022orca} and refined by
vLLM~\cite{kwon2023vllm} and
Sarathi-Serve~\cite{agrawal2024sarathi}. Continuous batching
dynamically recomposes the running batch at every decoder
iteration: requests that have just finished are evicted, and
queued requests can join mid-flight without waiting for the
entire batch to drain. vLLM's PagedAttention~\cite{kwon2023vllm}
additionally fragments the KV cache into pages of contiguous
tokens, allowing GPU memory to absorb more concurrent sequences
without external fragmentation.

These engine-level optimizations have transformed single-server
LLM serving but \textbf{only operate within one server}. The
question of \emph{which server should serve a given
request}---the focus of this paper---is left to an upstream
scheduler.

\subsection{Cloud-Edge Collaborative LLM Deployment}
\label{sec:bg-cloudedge}

Production LLM services increasingly deploy across
heterogeneous compute pools spanning cloud and edge
infrastructure. A typical topology includes:

\begin{itemize}
\item \textbf{One or more cloud servers} equipped with
  high-end accelerators (e.g., A100/H100, 80+~GB VRAM) capable
  of hosting the largest models (LLaMA-70B INT8 $\approx$ 70~GB)
  and serving high-throughput batches.

\item \textbf{A pool of heterogeneous edge servers} ranging
  from data-center A100 nodes through workstation-class T4 GPUs
  to embedded Jetson Orin devices. Edge nodes have smaller VRAM
  (16--64~GB), lower compute (10--50 TFLOPS), and lower power
  draw (30--250~W TDP). Smaller LLaMA-7B/13B variants can
  comfortably reside on mid-tier edge nodes; larger models
  cannot.

\item \textbf{A network fabric} with cloud-edge round-trip
  latency of 30--65~ms and edge-edge latency of 8--23~ms,
  modeled after typical regional WAN and metropolitan LAN
  deployments.
\end{itemize}

The cloud-edge paradigm is motivated by three economic forces:
(i)~\textbf{latency}---edge servers reduce first-token latency
for geographically proximate users; (ii)~\textbf{cost}---edge
GPUs are 1--10$\times$ cheaper per FLOP and 2--10$\times$ lower
power than cloud A100/H100 instances; (iii)~\textbf{data
sovereignty}---many enterprises restrict prompt data from
leaving local infrastructure. Splitwise~\cite{patel2024splitwise}
and DistServe~\cite{zhong2024distserve} have recently argued
for disaggregating prefill from decode across machines of
different generations to exploit hardware heterogeneity; we
share their premise but explore the \emph{scheduling} question
they leave open: given a fixed heterogeneous pool, where
should each request go?

\subsection{Scheduling Approaches: A Brief Survey}
\label{sec:bg-sched}

Cloud-edge scheduling has historically been addressed by three
families of algorithms, none of which natively encode AIGC
physics:

\textbf{Heuristic schedulers.} HEFT (Heterogeneous Earliest
Finish Time)~\cite{topcuoglu2002heft} computes a per-task
upward-rank priority and assigns each task to the server
minimizing estimated EFT under a fixed cost model. Round-Robin,
Least-Loaded, and Shortest-Queue are simpler load-balancing
heuristics popular in industry. These approaches treat tasks
as independent black boxes; they cannot, for example, prefer
a server that has the requested model already loaded.

\textbf{Metaheuristic search.} Genetic
Algorithms~\cite{holland1975adaptation} and Particle Swarm
Optimization~\cite{kennedy1995pso} perform population-based
search over candidate assignments under a designer-specified
fitness function. They are effective when the fitness landscape
is well-shaped but suffer from manual tuning of the fitness
function---and adding ``AIGC physics'' would require
hand-engineering additional fitness terms, which is precisely
the design we automate.

\textbf{Learning-based schedulers.} Decima~\cite{mao2019decima}
proposes GNN-based RL for Spark DAG scheduling, demonstrating
that neural policies can outperform hand-tuned heuristics under
realistic workloads. RLScheduler~\cite{zhang2020rlscheduler}
applies similar ideas to HPC batch jobs. Most relevantly to our
work, A3C-R2N2~\cite{tuli2022a3cr2n} uses asynchronous
actor-critic with a residual recurrent encoder for
\emph{generic} edge-cloud task scheduling---but it predates the
LLM era and includes no AIGC priors. Our scheduler builds on
this lineage but introduces AIGC-aware state and reward
components and contrasts directly against an A3C-R2N2 baseline
to isolate the AIGC contribution.

A complementary line of work focuses on within-server LLM
optimization rather than cross-server scheduling:
Splitwise~\cite{patel2024splitwise} disaggregates prefill and
decode onto different GPU generations;
DistServe~\cite{zhong2024distserve} co-optimizes resource
allocation and parallelism; Llumnix~\cite{sun2024llumnix}
performs runtime request migration across model instances.
These efforts are orthogonal to scheduler design---they could
be deployed inside any of the server pools our scheduler
manages---and we treat their batching and migration assumptions
as the substrate on which our scheduler operates.
