% =============================================================
% §3 Problem Formulation
% =============================================================
\section{Problem Formulation}
\label{sec:formulation}

We now formalize the cloud-edge LLM inference scheduling problem.
\S\ref{sec:sys-model} introduces notation for the deployment
topology and workload. \S\ref{sec:physics-constraints} expresses
the five AIGC physics as constraints and cost functions on
scheduling decisions. \S\ref{sec:sched-problem} states our
scheduling problem as a multi-objective optimization.
\S\ref{sec:mdp} casts it as a Markov Decision Process for
reinforcement-learning treatment.

\subsection{System Model}
\label{sec:sys-model}

\textbf{Deployment topology.} A heterogeneous compute pool
$\mathcal{S} = \{s_0, s_1, \ldots, s_N\}$ where $s_0$ is a cloud
server and $s_1, \ldots, s_N$ are edge servers. Each server $s$
has fixed attributes: compute capacity $C_s$ (TFLOPS), VRAM
capacity $V_s$ (GB), uplink bandwidth $B_s$ (Mbps), idle and
peak power $P^{\text{idle}}_s$ and $P^{\max}_s$ (W). A network
function $\tau(s, s', x)$ returns the transfer time of $x$~GB
of data between servers $s$ and $s'$.

\textbf{Model catalog.} A set
$\mathcal{M} = \{m_1, \ldots, m_K\}$ of LLM variants. Each model
$m$ has weight footprint $W_m$ (GB), cold-load time
$C^{\text{cold}}_m$ (s), per-token KV-cache size $\kappa_m$
(MB/token), per-token decode floor latency $\phi_m$ (s/token),
and maximum batch size $B^{\max}_m$.

\textbf{Request workload.} A stream of inference requests
$\mathcal{R} = \{r_1, r_2, \ldots\}$ arrives over time according
to a Poisson process at rate $\lambda$. Each request
$r_i = (m_i, L^p_i, L^o_i, a_i)$ has a target model
$m_i \in \mathcal{M}$, prompt length $L^p_i$, expected output
length $L^o_i$, and arrival time $a_i$.

\textbf{Task decomposition.} Each request $r_i$ decomposes into
two tasks $(t^{\text{pre}}_i, t^{\text{dec}}_i)$ with a strict
dependency edge $t^{\text{pre}}_i \to t^{\text{dec}}_i$. The
prefill task carries workload
$w^{\text{pre}}_i \propto L^p_i$, the decode task carries
$w^{\text{dec}}_i \propto L^o_i$, and both reference the same
model $m_i$ and request id $i$. The inference workload is
therefore a forest of pairwise-independent two-node DAGs.

\subsection{AIGC Physics as Constraints}
\label{sec:physics-constraints}

The five AIGC physics from \S\ref{sec:intro} manifest as the
following constraints and cost terms governing how tasks
execute on servers.

\textbf{(P1) Model weight residency.} Server $s$ holds a state
$L_s(t) \subseteq \mathcal{M}$ representing the set of models
currently loaded in its VRAM at time $t$. A task targeting
model $m$ admitted to $s$ pays a cold-load penalty
$C^{\text{cold}}_m$ if $m \notin L_s(t)$. If loading $m$ would
exceed $V_s$, the simulator evicts unpinned models in LRU
order. Formally, the admission delay is:
\begin{equation}
  \delta^{\text{cold}}_{s,m,t} = C^{\text{cold}}_m \cdot
    \mathbb{1}[m \notin L_s(t)].
\end{equation}

\textbf{(P2) KV-cache locality.} For each decode task
$t^{\text{dec}}_i$ scheduled to server $s$ when its prefill
$t^{\text{pre}}_i$ ran on $s'$, an additional KV-migration
cost applies:
\begin{equation}
  \delta^{\text{kv}}_i = \tau\!\left(s', s,
    L^p_i \cdot \kappa_{m_i} / 1024\right).
\end{equation}
When $s = s'$, this cost is zero (KV cache is local).

\textbf{(P3) Continuous batching.} Let $b_{s,m,k}(t)$ denote
the number of running same-model same-kind tasks on server $s$
for model $m$, phase $k \in \{\text{pre}, \text{dec}\}$. A new
task admitted at time $t$ to $(s, m, k)$ experiences effective
execution time
\begin{equation}
  T^{\text{exec}}_{\text{eff}}
    = T^{\text{solo}} \cdot
      \bigl(1 + b_{s,m,k}(t) \cdot \beta_k\bigr),
\end{equation}
where $\beta_{\text{pre}} = 0.30$ and
$\beta_{\text{dec}} = 0.05$ are phase-specific overhead factors.
Admission requires $b_{s,m,k}(t) < B^{\max}_m$ (batch slot not
full).

\textbf{(P4) Memory-bandwidth floor.} The solo execution time
of a phase task is the maximum of its compute-bound time and
its memory-bound floor:
\begin{equation}
  T^{\text{solo}}
    = \max\!\left(\frac{w_t}{C_s},\;
                  \phi_{m_t} \cdot n_t\right),
\end{equation}
where $w_t$ is the task's workload (TFLOPS) and $n_t$ is its
token count (input for prefill, output for decode).

\textbf{(P5) Two-phase dependency.} The decode task
$t^{\text{dec}}_i$ cannot become ready until
$t^{\text{pre}}_i$ has fully completed. Combined with (P1) and
(P2), this means a decode dispatched to $s \neq s'$ sees no
benefit from prefill's KV cache---either it pays migration
cost or re-runs prefill.

\subsection{Scheduling Problem}
\label{sec:sched-problem}

A scheduling policy $\pi$ produces, for each ready task at each
time step, an assignment $\pi: t \mapsto s \in \mathcal{S}$.
Given a request stream $\mathcal{R}$, $\pi$ induces:

\begin{itemize}
\item Per-request \textbf{TTFT} (time-to-first-token):
  $\text{TTFT}_i = e^{\text{pre}}_i - a_i$, where
  $e^{\text{pre}}_i$ is the wall-clock completion time of the
  prefill task.

\item Per-request \textbf{TPOT} (time-per-output-token):
  $\text{TPOT}_i = (e^{\text{dec}}_i - s^{\text{dec}}_i) / L^o_i$.

\item \textbf{\slo{} attainment}:
  \begin{equation}
    \text{SLO}(\pi) = \frac{1}{|\mathcal{R}|} \sum_i
      \mathbb{1}\!\bigl[\text{TTFT}_i \le T^*_{\text{TTFT}}
        \land \text{TPOT}_i \le T^*_{\text{TPOT}}\bigr].
  \end{equation}

\item \textbf{Energy per token}:
  \begin{equation}
    E_{\text{tok}}(\pi) =
      \frac{\sum_s \int_0^T P_s(t)\,dt}{\sum_i L^o_i},
  \end{equation}
  where instantaneous power
  $P_s(t) = P^{\text{idle}}_s + (P^{\max}_s - P^{\text{idle}}_s)
  \cdot u_s(t)$ depends on per-server compute utilization
  $u_s(t)$.
\end{itemize}

Our scheduling problem is the \textbf{bi-objective program}:
\begin{equation}
  \max_\pi\; \text{SLO}(\pi)
  \qquad
  \min_\pi\; E_{\text{tok}}(\pi),
\end{equation}
subject to admission feasibility (compute, VRAM, batch-slot)
at every step. Because the two objectives can be in tension
(e.g., always-cloud maximizes per-request compute but
overshoots energy budget), we evaluate policies on the
$(\text{SLO}, E_{\text{tok}})$ plane and seek \textbf{Pareto
dominance}: $\pi$ dominates $\pi'$ if $\pi$ is no worse on both
axes and strictly better on at least one.

\subsection{MDP Formulation for RL}
\label{sec:mdp}

We treat per-task dispatch as a sequential decision process and
solve it with reinforcement learning. At each ready task
arrival, the agent observes a state $\mathbf{s}_t$ summarizing
the current task's properties and all servers' utilization,
loaded-model sets, and batch occupancies (detailed encoding in
\S\ref{sec:rl}). The agent emits an action
$a_t \in \mathcal{S}$ choosing which server to dispatch to; the
simulator applies the assignment, advances time until the next
ready task, and emits a scalar reward $r_t$. We define the
reward as a weighted sum of seven physical and AIGC-aware terms
(specified in \S\ref{sec:rl}) designed to provide
\textbf{dense per-decision} feedback rather than waiting for
episode-end \slo{}/energy outcomes---an essential design choice
given the long episode horizons ($\sim$100 task dispatches per
simulation run).

The MDP is \textbf{partially observable} in principle (the
simulator's internal state is richer than $\mathbf{s}_t$), but
the observable features capture all decision-relevant
information under the simulator's dynamics. The MDP is
\textbf{non-stationary within an episode} (server utilization
evolves as tasks complete) and \textbf{stationary across
episodes} (workload distribution is fixed). We solve with PPO
with action masking, pretraining, and GAE, as detailed in
\S\ref{sec:rl}.

\textbf{Why RL.} A natural alternative to RL is solving the
integer-linear program implied by \S\ref{sec:sched-problem}
directly. We rejected this because (i) AIGC physics (P1)(P3)
introduce \emph{state-dependent} costs (cold-load and batch
overhead depend on server's running set, which changes within
an episode), breaking the static-cost ILP assumption; and
(ii) the combinatorial action space at each step
($|\mathcal{S}|^{|\mathcal{R}|}$) is too large for
branch-and-bound at realistic workload sizes. Heuristic and
metaheuristic schedulers (HEFT, GA, PSO) compose a more
comparable design space and serve as our primary baselines in
\S\ref{sec:eval}.
