% =============================================================
% §6 Discussion
% =============================================================
\section{Discussion}
\label{sec:discuss}

This section discusses what our findings mean for AIGC
scheduling research and practice
(\S\ref{sec:implications}), the limitations and threats to
validity of our study (\S\ref{sec:limitations}), and the
deployment-time guidance our results provide
(\S\ref{sec:guidance}).

\subsection{Implications for AIGC Scheduling Research}
\label{sec:implications}

Our experiments support three implications that extend beyond
the specific schedulers compared in \S\ref{sec:eval}.

\textbf{Implication 1: Simulator physics dominate algorithmic
sophistication.} The largest scheduler-level effect in our
entire study is the $-83\%$ \slo{} collapse from disabling
continuous batching simulation
(Table~\ref{tab:ablation}, Fig.~\ref{fig:ablation}). This
effect is roughly $10\times$ larger than any individual
reward-design or state-engineering choice we measured. The
implication is that AIGC scheduling researchers should invest
the bulk of their effort in faithful physical
simulation---particularly continuous batching, KV-cache
locality, and memory-bandwidth floors---rather than in clever
neural-network architectures or reward formulations. A
simulator that under-models AIGC physics will reward bad
scheduling policies; conversely, a physically faithful
simulator pulls even simple schedulers toward near-optimal
behavior.

\textbf{Implication 2: Pretraining is the cheapest essential
investment.} Disabling pretraining cost us $38\%$ \slo{} (the
second-largest ablation effect, $p<0.001$). Yet pretraining
requires only 20 episodes ($\sim$10--15 seconds of wall-clock)
on the target workload. This is essentially a ``free''
optimization that practitioners deploying RL schedulers should
always include. Importantly, pretraining matters not for
sample efficiency in the conventional sense (our policy
network is small and converges quickly), but for
\textbf{escaping locally optimal regions} of the policy
space---analogous to the ``always-cloud'' attractor we
documented in our diagnostic process that earlier
configurations exhibited before we introduced both extended
pretraining and the cloud-overuse penalty.

\textbf{Implication 3: Don't over-attribute gains to reward
shaping.} Our 12-component ablation revealed that no
individual AIGC reward component (warm, batch, affinity,
cloud-overuse, or even all jointly) significantly improves
\slo{} once continuous batching simulation and pretraining are
in place. We hypothesize that prior AIGC scheduling works
claiming specific gains from individual reward components
have likely overlooked this \textbf{substitutability} of
reward signals within a sufficiently expressive PPO backbone.
Future works should ablate aggressively: a positive result
from a single reward term is \emph{suspicious} unless backed
by the kind of joint ablation we report in
Table~\ref{tab:ablation}.

These three implications collectively suggest that the AIGC
scheduling research agenda should reorient from ``engineering
better rewards'' to \textbf{``engineering better simulators
and training procedures.''} The actual policy
network---whether MLP, GNN, or attention-based---matters less
than the environment in which it is trained.

\subsection{Limitations and Threats to Validity}
\label{sec:limitations}

Several limitations qualify our findings.

\textbf{Simulation fidelity.} Our simulator approximates
continuous batching as \emph{admission-time} batching (batch
size determined when a task is admitted, fixed thereafter),
rather than the iteration-level dynamic re-batching used by
vLLM. This simplification preserves the
throughput-vs-latency trade-off qualitatively but may
underestimate true continuous batching's benefit by 10--20\%
in some regimes. Our memory-bandwidth floor is calibrated
from published vLLM numbers but treats per-token latency as a
constant rather than batch-size-dependent (real HBM
contention scales sub-linearly with batch). We believe
neither simplification changes the qualitative findings, but
quantitative numbers (e.g., $-83\%$ from \texttt{no\_batching})
should be interpreted as our simulator's behavior rather than
vLLM-precise predictions.

\textbf{Workload coverage.} Our experiments use synthetic
log-normal prompt/output length distributions calibrated to
Azure LLM Inference Trace 2023 statistics. We have not
replayed true workload traces (which would require access to
proprietary production logs) nor tested on bursty arrival
patterns common in interactive workloads. The
\S\ref{sec:load} saturation finding (RL loses \slo{} edge at
$\lambda \ge 4$) might shift under bursty workloads where
short-term overload windows favor schedulers with explicit
batching control.

\textbf{Hardware coverage.} Our cloud-edge topology spans
A100/T4/Jetson-tier servers, matching published deployment
configurations. We have not tested with H100, MI300X, or
custom inference accelerators (e.g., Groq LPU). The
memory-bandwidth floor and batching constants would need
re-calibration for these. However, the qualitative AIGC
physics we model---weight residency, KV locality, batching,
bandwidth floor, two-phase tasks---are architecture-agnostic
and should generalize.

\textbf{Statistical power.} Our $N=30$ runs per configuration
give good detection power for medium-to-large effects
(Cohen's $d \ge 0.5$) but cannot resolve small effects
($d < 0.2$). The ablation null results in
\S\ref{sec:ablation} thus mean ``we cannot detect differences
larger than $\approx$5\% in \slo'' rather than ``no difference
exists.'' A more definitive ablation would require $N=100+$
runs, which we view as future work.

\textbf{Single learning algorithm.} Our scheduler uses PPO
with specific architectural choices (3-layer MLP, action mask,
GAE). Other RL families---offline RL, model-based RL,
imitation learning from a heuristic oracle---might respond
differently to AIGC reward shaping. We tested two
alternatives (A3C-R2N2 and GNN-PPO) and both exhibited similar
substitutability of AIGC components, suggesting our finding is
not PPO-specific, but the claim is bounded by this evidence.

\subsection{Practical Deployment Guidance}
\label{sec:guidance}

For practitioners deploying AIGC schedulers in production
cloud-edge environments, our results suggest the following
concrete guidance.

\textbf{Capacity planning matters more than scheduler choice.}
The \S\ref{sec:workload} large-heavy-model experiment shows
that when workload exceeds cluster capacity at a specific
tier (e.g., 70B requests overflowing the single cloud server),
all schedulers converge to $\approx$10\% \slo. A 28\% \slo{}
improvement from AIGC-aware scheduling is dwarfed by the
$3\times$ \slo{} improvement from adding a second cloud GPU.
\emph{Provision compute first; optimize scheduling second.}

\textbf{Use simple load balancers in the moderate-load regime}
($\lambda \le 2$ req/s with diverse model mix). Our
Fig.~\ref{fig:pareto} shows that LeastLoaded and ShortestQueue
achieve 78\% of RL's \slo{} attainment with 0\% of the
training cost. For deployments with small operational scale
or limited ML engineering capacity, the marginal benefit of
RL does not justify its operational overhead.

\textbf{Reach for RL when energy efficiency or tail latency
matter.} Across all 30 tested configurations, our AIGC-aware
RL achieved 5--11\% lower energy per token than the next-best
baseline (\S\ref{sec:topology}, \S\ref{sec:load}). For
large-scale deployments where a 5\% energy reduction
translates to meaningful annual savings (millions of USD at
hyperscaler volumes), the RL operational investment is
justified. Similarly, RL's stronger tail TTFT behavior (P95
25.2~s vs.\ 28--41~s for baselines at edge${=}5$,
$\lambda{=}2$) matters for interactive applications.

\textbf{Be skeptical of reward-engineering claims.} If a
future paper proposes a new AIGC reward component (X-bonus,
Y-penalty), ask: was it ablated jointly with other AIGC
rewards? Was the comparison against a pretrained baseline?
Our results suggest that without these controls, gains
attributed to specific reward components may reflect general
RL backbone improvements rather than the claimed mechanism.
