% =============================================================
% §7 Related Work
% =============================================================
\section{Related Work}
\label{sec:related}

Our work intersects four established lines of research:
(\S\ref{sec:rel-serving}) LLM serving systems,
(\S\ref{sec:rel-cluster}) cluster scheduling,
(\S\ref{sec:rel-rl}) reinforcement learning for scheduling,
and (\S\ref{sec:rel-aigc}) AIGC-specific scheduling. We
additionally discuss (\S\ref{sec:rel-sim}) simulation
platforms relevant to AIGC scheduling research.

\subsection{LLM Serving Systems}
\label{sec:rel-serving}

A wave of recent systems has dramatically improved
single-server LLM serving efficiency.
\textbf{Orca}~\cite{yu2022orca} introduced iteration-level
continuous batching, allowing requests to join and leave a
running batch at any decoder step rather than waiting for
batch completion. \textbf{vLLM}~\cite{kwon2023vllm} extended
this with \textbf{PagedAttention}, fragmenting the KV cache
into fixed-size pages that can be allocated non-contiguously,
eliminating internal fragmentation and enabling $2$--$4\times$
throughput improvements.
\textbf{Sarathi-Serve}~\cite{agrawal2024sarathi} introduced
\emph{chunked prefill}---decomposing long prefill phases into
multiple shorter chunks---to mitigate generation stalls in
mixed prefill/decode batches. NVIDIA's TensorRT-LLM provides
production-grade kernels with similar batching capabilities.

These systems are \textbf{orthogonal to our work}: they
optimize within a single server's GPU, whereas we schedule
requests across a heterogeneous cloud-edge pool. Our
simulator's admission-time batching model is a tractable
abstraction of their iteration-level batching; cross-server
scheduling decisions made by our RL would propagate into
vLLM-like engines at deployment time without architectural
change.

\subsection{Cluster Scheduling and Heuristics}
\label{sec:rel-cluster}

The cluster-scheduling literature predating LLM workloads
spans heuristic and metaheuristic approaches.
\textbf{HEFT}~\cite{topcuoglu2002heft} remains the canonical
static heuristic, computing upward-rank task priorities and
assigning each task to the server minimizing earliest finish
time. \textbf{Genetic
Algorithms}~\cite{holland1975adaptation} and \textbf{Particle
Swarm Optimization}~\cite{kennedy1995pso} are population-based
metaheuristics widely applied to scheduling fitness
landscapes. Industry deployments typically use simpler
load-balancing primitives (Round-Robin, Least-Loaded,
Shortest-Queue) for their operational predictability.

We compare against all of these in \S\ref{sec:eval}. None
natively encodes AIGC physics; adapting them would require
hand-engineering of AIGC-aware cost functions---an exercise
that motivates our learning-based approach.

\subsection{Reinforcement Learning for Scheduling}
\label{sec:rel-rl}

DRL-based scheduling has matured through several
representative works. \textbf{Decima}~\cite{mao2019decima}
uses a graph neural network over Spark DAG structure with
REINFORCE-style training, demonstrating that RL can learn
workload-specific policies outperforming hand-tuned
heuristics.
\textbf{RLScheduler}~\cite{zhang2020rlscheduler} applies
kernel-based RL to HPC batch jobs.
\textbf{A3C-R2N2}~\cite{tuli2022a3cr2n} uses asynchronous
actor-critic with a residual recurrent encoder for edge-cloud
IoT task scheduling---the closest predecessor to our work in
deployment topology, though predating the AIGC era.

We position our RL scheduler against this lineage by
\textbf{directly benchmarking A3C-R2N2 as a controlled
baseline} in \S\ref{sec:eval}. Holding the actor-critic
algorithm constant, the only differences are (i) our
AIGC-aware state encoding and (ii) our AIGC-aware reward
shaping. The Pareto improvement we measure
(\S\ref{sec:main-results}: \slo{} $+28\%$, energy $-7\%$, both
$p<0.05$) is thus attributable to AIGC awareness specifically,
not to general RL machinery. We also implement a \textbf{GNN
variant} inspired by Decima's encoder structure, finding that
under realistic training budgets a feedforward MLP with
explicit AIGC features matches or exceeds the GNN's
relational reasoning---an interesting departure from Decima's
original Spark-DAG findings.

\subsection{AIGC-Specific Scheduling}
\label{sec:rel-aigc}

The newest research wave addresses LLM-specific scheduling
properties directly. \textbf{Splitwise}~\cite{patel2024splitwise}
proposes \textbf{phase splitting}: routing prefill to
compute-rich GPUs (H100) and decode to memory-rich GPUs
(A100), with explicit KV-cache transfer across the
disaggregation boundary.
\textbf{DistServe}~\cite{zhong2024distserve} co-optimizes
resource allocation and parallelism strategy across
disaggregated prefill/decode pools.
\textbf{Llumnix}~\cite{sun2024llumnix} performs runtime
request \emph{migration} across model instances to address
request heterogeneity, treating each instance as a ``CPU
core'' in an operating-system-style scheduling abstraction.

These works share our motivation---LLM workloads need
LLM-aware scheduling---but address a different problem layer:
they design \emph{deployment topologies} (which GPUs serve
which phase) and \emph{intra-cluster reactive policies}
(mid-flight migration), assuming a fixed topology and
treating \emph{initial request placement} as orthogonal.
\textbf{Our work attacks the initial placement problem}:
given a fixed heterogeneous pool, learn a per-request
server-selection policy that co-optimizes \slo{} attainment
and energy efficiency. The two strands compose: Llumnix-style
migration could refine our scheduler's initial placement at
runtime, and our scheduler could populate prefill/decode
pools in a Splitwise-style deployment.

\subsection{Simulation Platforms}
\label{sec:rel-sim}

Several open-source simulators have supported scheduling
research: \textbf{CloudSim}~\cite{calheiros2011cloudsim} and
\textbf{iFogSim}~\cite{gupta2017ifogsim} for general
cloud-fog workloads;
\textbf{WorkflowSim}~\cite{chen2012workflowsim} for DAG
workflows. None of these models the AIGC-specific physics
central to our work (model weights, KV cache, continuous
batching, memory-bandwidth floors). A handful of recent
open-source projects target LLM serving simulation, but to
our knowledge they focus on intra-server batching dynamics
(modeling the behavior of a single inference engine) rather
than cross-server scheduling across heterogeneous pools.

To our knowledge, our simulator (\S\ref{sec:simulator}) is the
first open-source platform combining three properties:
(i)~modeling all five AIGC physics listed in \S\ref{sec:intro}
(model-weight residency, KV-cache locality, continuous
batching, memory-bandwidth floor, two-phase request
lifecycle); (ii)~focusing on cross-server scheduling across
heterogeneous pools rather than intra-server batching
dynamics; (iii)~supporting multi-objective (\slo{}~+~energy)
evaluation natively rather than as a post-hoc analysis. We
release it with reproducibility manifests for the 30+
configurations reported in \S\ref{sec:eval}, hoping to lower
the activation energy for future AIGC scheduling research.
