% =============================================================
% §8 Conclusion
% =============================================================
\section{Conclusion}
\label{sec:conclusion}

This paper investigated AIGC-aware reinforcement-learning
scheduling for cloud-edge LLM inference. We built a
comprehensive AIGC simulator modeling five physical
characteristics---model weight residency, KV-cache locality,
continuous batching, memory-bandwidth floors, and two-phase
request lifecycle---then evaluated nine schedulers spanning
load-balancing heuristics, classic and metaheuristic DAG
schedulers, and three RL variants (PPO with MLP,
A3C-R2N2~\cite{tuli2022a3cr2n} with GRU encoder, and PPO with
a Decima-inspired~\cite{mao2019decima} GNN encoder).

Our AIGC-aware PPO scheduler achieves \textbf{Pareto
dominance} over eight strong baselines on the (\slo{}
attainment, energy per token) plane: $+28\%$ \slo{} and
$-7\%$ energy at the canonical edge${=}5$, $\lambda{=}2$
configuration, significant at $p<0.05$ on \slo{} against all
four strongest comparators and on energy efficiency against
three of four (PSO $p=0.11$, others $p<0.05$). The Pareto
dominance extends robustly across edge counts $\{3, 5, 7\}$
and survives the strict A3C-R2N2 controlled comparison that
isolates the AIGC contribution from general RL machinery.

A 12-component ablation, however, reveals that \textbf{AIGC-aware
scheduling is a holistic system contribution rather than a
clever reward trick}: only continuous batching simulation
($-83\%$ \slo{} ablated) and adequate pretraining ($-38\%$
\slo{} ablated) emerge as significant components. All
individual AIGC reward and state components are statistically
indistinguishable from the Full RL baseline in isolation.
This finding cautions the community against claiming gains
from specific AIGC reward terms without joint ablation.

We hope this work supports two longer-term directions: more
\textbf{realistic AIGC scheduling benchmarks} built on the
open platform we release, and a \textbf{research culture of
aggressive joint ablation} for AIGC scheduling claims. The
fast-moving LLM-systems landscape will reward both.
