% =============================================================
% 中英文摘要（中文期刊惯例：双语并存）
% =============================================================

\begin{abstract}
\textbf{中文摘要。}大语言模型（LLM）推理负载的快速增长，给数据中心带来巨大的服务成本与能耗压力，仅以吞吐量为单一优化目标的传统调度器已难以胜任。本文研究\textbf{云--边协同的 LLM 推理调度问题}，并显式建模五项 AIGC 特有的物理约束：模型权重驻留（7--70~GB）、KV 缓存随 token 增长的显存占用、连续批处理（continuous batching）、解码阶段的显存带宽下限，以及预填充--解码（prefill/decode）两阶段请求生命周期。我们设计了一个面向 AIGC 物理的 PPO 调度器，将上述物理量通过状态特征与奖励整形显式暴露给策略网络，并在涵盖启发式（HEFT、GA、PSO）、负载均衡（LeastLoaded、ShortestQueue）与学习式（A3C-R2N2、GNN）的 8 个强基线上做对比。在 30 种配置（拓扑边缘数 $\in\{3,5,7\}$、到达率 $\lambda \in \{0.5,1,2,4,8\}$ 请求/秒、工作负载组合分别为 uniform 与 large-heavy）下，每种配置 $N=30$ 次独立运行。在典型工况（$\lambda{=}2$ 请求/秒、uniform 混合、edge${=}5$）下，本文调度器同时达到 \slo{} 占优与能效占优——相对最强基线，\slo{} 达成率提升 28\%，每 token 能耗降低 7\%，\slo{} 维度对全部 4 个强对手 $p<0.05$ 显著，能效维度对其中 3 个显著（PSO $p=0.11$）。然而 12 组件消融分析揭示出一个反直觉发现：联合改进是\textbf{系统级整体贡献}而非某项奖励技巧——仅"连续批处理仿真"与"充足预训练"个体显著，所有单项 AIGC 状态或奖励组件在单独消融时与完整模型在统计上不可区分。我们开源仿真器及 30+ 套基准 manifest，以支持后续 AIGC 调度研究。

\textbf{关键词：}LLM 推理；AIGC；云边协同；强化学习；调度；能效；Pareto 优化。

\textbf{中图分类号：}TP18

\textbf{文献标识码：}A
\end{abstract}

% --- English Abstract（中文期刊要求双语并存） -----------------
\noindent\textbf{Abstract.} The rapid growth of large language
model (LLM) inference workloads imposes serving-cost and energy
pressures that single-axis throughput-oriented schedulers cannot
address. We study \emph{cloud--edge collaborative LLM inference
scheduling} under five AIGC-specific physical constraints:
model-weight residency (7--70~GB), KV-cache memory growth,
continuous batching, the memory-bandwidth floor of decode, and
the two-phase prefill/decode request lifecycle. We design an
AIGC-aware proximal policy optimization (PPO) scheduler that
exposes these physics to the policy through state features and
reward shaping, and compare against eight strong baselines
spanning heuristic (HEFT, GA, PSO), load-balancing (LeastLoaded,
ShortestQueue), and learned (A3C-R2N2, GNN) approaches. Across
30 configurations (edge $\in \{3,5,7\}$,
$\lambda \in \{0.5,1,2,4,8\}$ req/s, uniform vs.\ large-heavy
mixes) with $N{=}30$ runs each, our scheduler achieves strict
Pareto dominance in the canonical operating regime
($\lambda{=}2$~req/s, uniform mix, edge${=}5$): $+28\%$ \slo{}
attainment and $-7\%$ energy per token, with significance at
$p<0.05$ on \slo{} against all four strong comparators and on
energy efficiency against three of four. A 12-component
ablation reveals that the gain is a holistic system
contribution: only continuous-batching simulation and adequate
pretraining are individually significant; every AIGC reward and
state component is statistically indistinguishable from the
full model when ablated alone. We release the simulator and the
30+ benchmark manifests to support future AIGC scheduling
research.

\noindent\textbf{Key words:} LLM inference; AIGC; cloud-edge
computing; reinforcement learning; scheduling; energy
efficiency; Pareto optimization.
